Digging Deep to Assess the State of the Practice for Different Research Software Domains

### 2022-02-15 StateOfPractice/Papers/ICCS-2022-SOP_MethodologyPlusExamples/DigDeep.tex
Updates to SOP for LBM, start of digging deeper paper

\documentclass[letterpaper,cleveref]{lipics-v2019}

\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage{paralist}
\usepackage{totcount}
\usepackage{booktabs}

\hypersetup{
colorlinks=true,     % false: boxed links; true: colored links
linkcolor=blue,      % color of internal links (change box color with linkbordercolor)
citecolor=blue,      % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan        % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{brown}{AD}{#1}} %Ao Dong

\newcommand{\notdone}[1]{\textcolor{red}{#1}}
\newcommand{\done}[1]{\textcolor{black}{#1}}

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm

\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf
+}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}}

@@

\title{Methodology for Assessing the State of the Practice for Domain X} 
\author{Spencer Smith}{McMaster University, Canada}{smiths@mcmaster.ca}{}{}
\author{Jacques Carette}{McMaster University, Canada}{carette@mcmaster.ca}{}{}
\author{Peter Michalski}{McMaster University, Canada}{michap@mcmaster.ca}{}{}
\author{Ao Dong}{McMaster University, Canada}{donga9@mcmaster.ca}{}{}
\author{Olu Owojaiye}{McMaster University, Canada}{owojaiyo@mcmaster.ca}{}{}

\authorrunning{Smith et al.}  \Copyright{Spencer Smith and Jacques Carette and
 Peter Michalski and Ao Dong and Olu Owojaiye}

\ccsdesc{Software and its engineering}
\ccsdesc{Software and its engineering~Software product lines}
\ccsdesc{General and reference~Empirical studies}

\keywords{software quality, domain analysis, scientific computing software,
  research software, empirical measures, analytic hierarchy process}

\date{\today}

\hideLIPIcs
\nolinenumbers

\begin{document}

\maketitle

\begin{abstract}
	To improve software development methods and tools for research software, we
	first need to understand the current state of the practice.  Therefore, we
	have developed a methodology for assessing the state of the software
	development practices for a given research software domain.  The methodology
	is applied to one domain at a time in recognition that software development in
	different domains is likely to have adopted different best practices.
	Moreover, providing a means to measure different domains facilitates
	comparison of development practices between domains.  For each domain we wish
	to answer questions such as: 
  \begin{inparaenum}[i)]
    \item What artifacts (documents, code, test cases, etc.) are present?
    \item What tools are used?
    \item What principles, process and methodologies are used?
    \item What are the pain points for developers?
    \item What actions are used to improve qualities like maintainability and
    reproducibility?
  \end{inparaenum} 
  To answer these questions, our methodology prescribes the following steps: 
  \begin{inparaenum}[i)] 
    \item Identify the domain;
    \item Identify a list of candidate software packages;
    \item Filter the list to a length of about 30 packages;
    \item Gather source code and documentation for each package;
    \item Collect repository related data on each software package, like number
    of stars, number of open issues, number of lines of code;
    \item Fill in the measurement template (the template consists of 108
    questions to assess 9 qualities (including the qualities of installability,
    usability and visibility));
    \item Interview developers (the interview consists of 20 questions and takes
    about an hour);
    \item Rank the software using the Analytic Hierarchy Process (AHP); and,
    \item Analyze the data to answer the questions posed above.
  \end{inparaenum}
  A domain expert should be engaged throughout the process, to ensure that
  implicit information about the domain is properly represented and to assist
  with conducting an analysis of the commonalities and variabilities between the
  30 selected packages.  Using our methodology, spreadsheet templates and AHP
  tool, we estimate (based on our experience with using the process) the time to
  complete an assessment for a given domain at \total{totHours} person hours.
\end{abstract}

~\newpage

\tableofcontents

~\newpage

\section{Introduction} \label{SecIntroduction}

@@

Given the importance of research software, scientists and engineers are pushing
for methods and tools to sustainably develop high quality software.  This is
evident from the existence of such groups as the Software Sustainability
Institute (\href{https://www.software.ac.uk/} {SEI}) and Better Scientific
Software (\href{https://bssw.io/} {BSS}).  Sustainability promoting groups such
as these are necessary because unfortunately the current ``state of the
practice'' for research software often does not incorporate ``state of the art''
Software Engineering (SE) tools and methods \citep{JohansonAndHasselbring2018}.
The lack of SE tools and methods contributes to sustainability and reliability
problems \citep{FaulkEtAl2009}.  Problems with the current state of the practice
are evident from embarrassing failures, like a retraction of derived molecular
protein structures \citep{Miller2006}, false reproduction of sonoluminescent
fusion \citep{PostAndVotta2005}, and fixing and then reintroducing the same
error in a large code base three times in 20 year
\citep{MilewiczAndRaybourn2018}.  To improve this situation, we need to first
fully understand the current state of the practice for research software
development.

@@

This ``state of the practice for domain X'' exercise builds off of prior work on
measuring/assessing the state of software development practice in several
research domains.  We have updated the work that was done previously for domains
such as Geographic Information Systems \citep{SmithEtAl2018_arXivGIS}, Mesh
Generators \citep{SmithEtAl2016}, Seismology software \citep{SmithEtAl2018}, and
Statistical software for psychology \citep{SmithEtAl2018_StatSoft}.  Initial
tests of the new methodology have been done for medical image analysis software
\citep{Dong2021} and for Lattice Boltzmann Method (LBM) software
\citep{Michalski2021}.

@@

In the proposed methodology, the collected data is combined to rank the software
within each domain using the Analytic Hierarchy Process (AHP) \citep{Saaty1980}.
As in the previous measurement exercise, we use AHP to develop a list of
software ranked by quality.  However, in the new process we do not stop with
this list.  The Domain Expert is consulted to verify the ordering, and to
discuss the decisions that led to the ranking.  The AHP process is used to
facilitate a conversation with the Domain Expert as a means to deepen our
understanding of the software in the domain, and the needs of typical
developers.

@@

The scope of this methodology includes observations on product, artifact
(documentation, test scripts, etc.) and process quality for research software.
We leave the assessment of the performance of research software, for instance
using benchmarks, to other projects, such as the work of
\citet{kaagstrom1998gemm}.  Currently we are also leaving experiments to measure
usability and modifiability as future work, as discussed in
Section~\ref{SecFutureWork}.

Within this document the following typographic conventions have been adopted:
\begin{inparaenum}[i)] 
  \item \textcolor{blue}{Blue} text denotes a link to sections of this document,
  including citations 
  \item \textcolor{cyan}{Cyan} text denotes an external URL link.
\end{inparaenum}

@@

A domain of research software must be identified to begin the assessment.
Research software is defined in this exercise as ``software that is used to
generate, process or analyze results that [are intended] to appear in a
publication'' \citep{hettrick2014uk} in a scientific or engineering context.
Research software is a more general term for what is often called scientific
computing.  To be applicable for the methodology described in this document, the
chosen domain must have the following properties:

@@

Some examples of domains that fit these criteria are finite element analysis
\citep{szabo1996finite}, quantum chemistry \citep{veryazov20042molcas},
seismology \citep{SmithEtAl2018}, as well as mesh generators
\citep{smith2016state}.

@@

Since each domain we will study will have a reasonably small scope, we will be
able to view the software as constituting a program family.  The concept of a
program family is defined by \citet{parnas1976design} as ``a set of programs
whose common properties are so extensive that it is advantageous to study the
common properties of the programs before analyzing individual members''.
Studying the common properties within a family of related programs is termed a
domain analysis.

@@

\begin{figure}[!ht]
	\begin{center}
	  \includegraphics[width=1.0\textwidth]{measurement_template}
	  \caption{Excerpt of the Top Section of the Measurement Template (Summary Information)}
	  \label{measurement_template_image}
	\end{center}
  \end{figure}

The full template consists of 108 questions categorized under 9 qualities.  The
questions were designed to be unambiguous, quantifiable and measurable with
limited time and domain knowledge. The measures are grouped under headings for
each quality, and one for summary information. The summary information (shown in
Figure~\ref{measurement_template_image}) is the first section of the template.
This section summarizes general information, such as the software name, number
of developers, etc.  We follow the definitions given by
\citet{GewaltigAndCannon2012} for the software categories.  Public means software
intended for public use.  Private means software aimed only at a specific
group, while the concept category is used for software written simply to demonstrate
algorithms or concepts. The three categories of development models are: open
source, where source code is freely available under an open source license;
free-ware, where a binary or executable is provided for free; and, commercial,
where the user must pay for the software product.

@@

As in \citet{SmithEtAl2016}, Virtual machines (VMs) are used to provide an
optimal testing environments for each package. VMs were used because it is
easier to start with a fresh environment without having to worry about existing
libraries and conflicts. Moreover, when the tests are complete the VM can be
deleted, without any impact on the host operating system. The most significant
advantage of using VMs is to level the playing field. Every software install
starts from a clean slate, which removes ``works-on-my-computer'' errors. When
filling in the measurement template spreadsheet, the the details for each VM
should be noted, including hypervisor and operating system version.

@@

In the previous state of the practice measurement process
\citep{SmithEtAl2018_arXivGIS, SmithEtAl2016, SmithEtAl2018}, we only based our
assessment on information available in on-line software repos.  However, this
approach meant we weren't able to learn about the development process, the
attitudes of the developers, the pain points for developers and how the
developers handle modifiability, reproducibility and usability.  Therefore, in
the reboot of the measurement process, we have explicitly added a stage for
interviewing research software developers.

@@

Table~\ref{TabPersonHours} estimates the time required (in person hours) to
complete a state of the practice assessment for domain X.  The table assumes
that the domain has already been decided and the Domain Expert has been
recruited.  The time spent by the Domain Expert is not included in the numbers
shown in the table, since the amount of time that the domain expert will work
independently of the rest of the assessment team will be small.  Moreover, this
amount of time will vary greatly depending on the preferred work habits of the
Domain Expert.  The table follows the steps outlined in Section~\ref{StepsAQDS}.
Time is not included for reviewing the methodology. Moreover, it is assumed that
the template spreadsheets linked in this document, and the developed AHP tool,
will be employed, rather than developing new tools.  The person hours given are
a rough estimate, based on our experience completing assessments for medical
image analysis software \citep{Dong2021} and for Lattice Boltzmann Method (LBM)
software \citep{Michalski2021}.  These two domains were assessed at the same
time as designing the methodology presented in this document.  We did our best
to estimate the time spent on measurement and separate it from the time spend on
design and development.  The estimate assumes 30 software packages will be
measured; the numbers will need to be adjusted if the total packages changes.

@@

As explained in the introduction (Section~\ref{SecIntroduction}), our eventual
goal is to improve the state of practice of software development, which requires
a baseline means for measuring the current state of the practice.  Now that we
have a methodology for assessing the state of practice for a given domain, the
next task is to complete the measurements for multiple domains.  Using the
previous and updated methodologies, we have measured the state of
practice for the following domains: Geographic Information Systems
\citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Seismology software \citep{SmithEtAl2018}, Statistical software for psychology
\citep{SmithEtAl2018_StatSoft}, medical image analysis software \citep{Dong2021}
and for Lattice Boltzmann Method software \citep{Michalski2021}.  Future domains
for measurement could include finite element software, computational medicine,
machine learning, ordinary differential equation solvers, computer graphics,
stoichiometry, etc.

@@

If we can add measures for modifiability and usability, we can start to measure
the quality impact of software development processes, tools and techniques.  For
instance, we have a project (called
\href{https://github.com/JacquesCarette/Drasil} {Drasil}
\citep{SzymczakEtAl2016}), which facilitates a development process that focuses
on knowledge capture, followed by generation of code and other software
artifacts from the captured knowledge.  To understand the advantages and
disadvantages of Drasil, we could measure its quality.  In particular, we would
like to see the impact of a generative process on the quality of usability and
modifiability.

@@

The next experiment is designed to gather qualitative data regarding the
modifiability of each software package. This proposed experiment also involves
experimental subjects/participants, who in this case are asked to modify a set
of software packages.  The specific modifications requested will again depend on
the software domain.  In advance of the experiment the Domain Expert will be
asked the likely changes for software in this domain.  We emphasize likely
changes, instead of any changes because software cannot be designed so that
everything is equally easy to change \citep{parnas1986rational}.  The procedure
could be along the following lines:

@@

Combining the quantitative data from the measurement template with the interview
results, along with the domain experts knowledge, we can determine the current
state of the practice for domain X.  Using our methodology, spreadsheet
templates and AHP tool, we estimate (based on our experience with using the
process) the time to complete an assessment for a given domain at
\total{totHours} person hours.

\newpage

\appendix

\section{A Guide to Repository-Based Measurement Tools} \label{SecRepoTools}

This appendix covers the tools used for collecting repository based data
(Section~\ref{SecEmpiricalMeasures}).  The two tools covered are git\_stats and
scc.  The tools do not have to be used in any particular order.

\subsection{git\_stats}

\subsubsection{Introduction}

Source Code: \href{https://github.com/tomgi/git_stats}{GitHub repo}

\subsubsection{User Manual} \label{git_stats_manual}

Official Manual: \href{https://github.com/tomgi/git_stats}{GitHub repo}

\subsubsection{Demo of Installation and Running the Tool}

The installation steps on your machine may be different from this section.
Please refer to the user manual mentioned in Section \ref{git_stats_manual}, if
necessary.  The steps shown here were executed on a virtual machine with 8 cores
and 16 GB RAM running Debian GNU/Linux 9.11.

\begin{enumerate}
\item Install ruby/gem environment
\begin{lstlisting}
apt-get install ruby ruby-nokogiri ruby-nokogiri-diff ruby-nokogumbo
\end{lstlisting}

Check the installation:
\begin{lstlisting}
gem --version
\end{lstlisting}

\item Install the tool
\begin{lstlisting}
sudo gem install git_stats
\end{lstlisting}

\item Prepare the target repo

Make sure the target repo (the repo to be analyzed, not the repo of this tool)
is on your machine.
In this demo, the target repo is downloaded from a GitHub repo:
\begin{lstlisting}
# change [git path] to the url of your target repo
git clone [git path]
# e.g. git clone https://github.com/nroduit/Weasis.git
\end{lstlisting}

\item Generate analytics
\begin{lstlisting}
# make sure [repo path] is the target repo path
# the [output path] can be anywhere you desire
git_stats generate -p [repo path] -o [output path]
# e.g. git_stats generate -p /home/user/git-stats/Weasis -o 
# /home/user/git-stats/Weasis-analytics
\end{lstlisting}

\item View the analytics

View the analytic results by open [output path]/index.html with any browser or
other software supporting HTML web page format.

\begin{figure}[!ht]
\includegraphics[scale=0.4]{git-stats-demo.JPG}
\end{figure}

\item Download the data

On most of the taps of this web page, the data can be downloaded for more
analytics by clicking the menu button beside the data-range section.

\end{enumerate}

\subsection{scc}

\subsubsection{Introduction}

Source Code: \href{https://github.com/boyter/scc}{GitHub repo}

\subsubsection{User Manual} \label{scc_manual}

Official Manual: \href{https://github.com/boyter/scc}{GitHub repo}

\subsubsection{Demo of Installation and Running the Tool}

The installation steps on your machine may be different from this section.
Please refer to the user manual mentioned in Section \ref{scc_manual}, if
necessary.  These steps were executed on a virtual machine with 8 cores and 16
GB RAM running Debian GNU/Linux 9.11

\begin{enumerate}
\item Install Golang

Follow the \href{https://golang.org/doc/install}{official instructions}, or the
following demo,

download the installation package:
\begin{lstlisting}
wget https://dl.google.com/go/go1.14.3.linux-amd64.tar.gz
\end{lstlisting}

unpack to /usr/local:
\begin{lstlisting}
sudo tar -C /usr/local -xzf go1.14.3.linux-amd64.tar.gz
\end{lstlisting}

use a text editor to open \textasciitilde/.profile, e.g.:
\begin{lstlisting}
nano ~/.profile
\end{lstlisting}

add the following lines to the end of this file:
\begin{lstlisting}
export GOPATH=$HOME/go
export PATH=$PATH:/usr/local/go/bin:$GOPATH/bin
\end{lstlisting}

save the file, and load the commands into the current shell instance:
\begin{lstlisting}
source ~/.profile
\end{lstlisting}

check the installation:
\begin{lstlisting}
go version
\end{lstlisting}

\item Install the tool
\begin{lstlisting}
go get -u github.com/boyter/scc/
\end{lstlisting}

\item Prepare the target repo

Make sure the target repo (the repo to be analyzed, not the repo of this tool)
is on your machine.
In this demo, the target repo is downloaded from a GitHub repo:
\begin{lstlisting}
# change [project path] to your desired folder
cd [project path]
git clone https://github.com/nroduit/Weasis.git
\end{lstlisting}

\item Generate analytics
\begin{lstlisting}
# make sure [repo path] is the target repo path
cd [repo path]
# use scc to generate analytics
scc
\end{lstlisting}

\item View the analytics

The results will be appear something like the following.

\begin{figure}[!ht]
\includegraphics[scale=0.5]{scc-demo.JPG}
\end{figure}

\end{enumerate}

\newpage

\section{Measurement Template} \label{SecGradingTemplate}

The table below lists the full set of measures that are assessed for each
software product.  The measures are grouped under headings for each quality, and
one for summary information.  Following each measure, the type for a valid
result is given in brackets.  Many of the types are given as enumerated sets.
For instance, the response on many of the questions is one of ``yes,'' ``no,''
or ``unclear.''  The type ``number'' means natural number, a positive integer.  The
types for date and url are not explicitly defined, but they are what one would
expect from their names.  In some cases the response for a given question is not
necessarily limited to one answer, such as the question on what platforms are
supported by the software product.  Case like this are indicated by ``set of''
preceding the type of an individual answer.  The type in these cases are then
the power set of the individual response type.  In some cases a superscript $^*$
is used to indicate that a response of this type should be accompanied by
explanatory text.  For instance, if problems were caused by uninstall, the
reviewer should note what problems were caused.  % An (I) precedes the question
% description when its measurement requires a successful installation.
The template also include 3 sections at the end for summarizing the repository
based metrics.  A
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}
{blank measurement template spreadsheet} is available to save time with data
entry.

\newpage

\captionof{table}{Measurement Template}\label{measurementtemplate}
\def\arraystretch{1.22}
\begin{tabular}{p{14cm}}
	\hline
	\textbf{Summary Information}\\
	\hline
	Software name? (string)\\
	URL? (URL)\\
	Affiliation (institution(s)) (string or {N/A})\\
	Software purpose (string)\\
	Number of developers (all developers that have contributed at least one commit
	to the project) (use repo commit logs) (number)\\
	How is the project funded? (unfunded, unclear, funded$^*$) where $^*$ requires a
	string to say the source of funding\\
	Initial release date? (date)\\
	Last commit date? (date)\\
	Status? (alive is defined as presence of commits in the last 18 months)
	({alive, dead, unclear})\\
	License? ({GNU GPL, BSD, MIT, terms of use, trial, none, unclear, other$^*$}) $^*$
	given via a string \\
	Platforms? (set of {Windows, Linux, OS X, Android, other$^*$}) $^*$ given via
	string\\
	Software Category? The concept category includes software that does not have
	an officially released version. Public software has a released version in the
	public domain. Private software has a released version available to authorized
	users only. ({concept, public, private})\\
	Development model? ({open source, freeware, commercial, unclear})\\
	Publications about the software? Refers to publications that have used or
	mentioned the software. (number or {unknown})\\
	Source code URL? ({set of url, n/a, unclear})\\
	Programming language(s)? (set of {FORTRAN, Matlab, C, C++, Java, R, Ruby,
	Python, Cython, BASIC, Pascal, IDL, unclear, other$^*$}) $^*$ given via string \\
	Is there evidence that performance was considered? Performance refers to
	either speed, storage, or throughput. ({yes$^*$, no})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have) \\
	\hline
\end{tabular}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
% \usepackage{color}
% \renewcommand\UrlFont{\color{blue}\rmfamily}
%
\usepackage{hyperref}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{totcount}

@@

\begin{document}
%
\title{Digging Deep to Assess the State of the Practice for Different Research Software Domains}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Spencer Smith\inst{1}\orcidID{0000-0002-0760-0987}} %\and
%Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{S.\ Smith}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{McMaster University, 1280 Main Street West, Hamilton ON L8S 4K1, Canada
\email{smiths@mcmaster.ca}\\
\url{http://www.cas.mcmaster.ca/~smiths/}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}

	To improve software development methods and tools for research software, we
	first need to understand the current state of the practice.  Therefore, we
	have developed a methodology for assessing the state of the software
	development practices for a given research software domain.  The methodology
	is applied to one domain at a time in recognition that software development in
	different domains is likely to have adopted different best practices.
	Moreover, providing a means to measure different domains facilitates
	comparison of development practices between domains.  For each domain we wish
	to answer questions such as: 
  \begin{inparaenum}[i)]
    \item What artifacts (documents, code, test cases, etc.) are present?
    \item What tools are used?
    \item What principles, process and methodologies are used?
    \item What are the pain points for developers?
    \item What actions are used to improve qualities like maintainability and
    reproducibility?
  \end{inparaenum} 
  To answer these questions, our methodology prescribes the following steps: 
  \begin{inparaenum}[i)] 
    \item Identify the domain;
    \item Identify a list of candidate software packages;
    \item Filter the list to a length of about 30 packages;
    \item Gather source code and documentation for each package;
    \item Collect repository related data on each software package, like number
    of stars, number of open issues, number of lines of code;
    \item Fill in the measurement template (the template consists of 108
    questions to assess 9 qualities (including the qualities of installability,
    usability and visibility));
    \item Interview developers (the interview consists of 20 questions and takes
    about an hour);
    \item Rank the software using the Analytic Hierarchy Process (AHP); and,
    \item Analyze the data to answer the questions posed above.
  \end{inparaenum}
  A domain expert should be engaged throughout the process, to ensure that
  implicit information about the domain is properly represented and to assist
  with conducting an analysis of the commonalities and variabilities between the
  30 selected packages.  Using our methodology, spreadsheet templates and AHP
  tool, we estimate (based on our experience with using the process) the time to
  complete an assessment for a given domain at 173 person hours.

The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction} \label{SecIntroduction}

- research software is vital for making progress in science and engineering, but it is difficult to develop
- this has promoted many studies to understand the challenges of research software development
- the previous studies have usually used interviews, surveys and case studies
- the interviews and surveys recruit participants from all domains of research software
- the case studies, by their nature, focus on specific examples
- what is missing is a methodology that:
	- goes beyond surveys to look at the actual artifacts in the software repos
	- focuses on one domain at a time
- by focusing on one domain at a time we can:
	- provide useful knowledge/advice to the developers in that domain
	- build up a set of measures of different domains to compare and contrast them

@@

Given the importance of research software, scientists and engineers are pushing
for methods and tools to sustainably develop high quality software.  This is
evident from the existence of such groups as the Software Sustainability
Institute (\href{https://www.software.ac.uk/} {SEI}) and Better Scientific
Software (\href{https://bssw.io/} {BSS}).  Sustainability promoting groups such
as these are necessary because unfortunately the current ``state of the
practice'' for research software often does not incorporate ``state of the art''
Software Engineering (SE) tools and methods \cite{JohansonAndHasselbring2018}.
The lack of SE tools and methods contributes to sustainability and reliability
problems \cite{FaulkEtAl2009}.  Problems with the current state of the practice
are evident from embarrassing failures, like a retraction of derived molecular
protein structures \cite{Miller2006}, false reproduction of sonoluminescent
fusion \cite{PostAndVotta2005}, and fixing and then reintroducing the same
error in a large code base three times in 20 year
\cite{MilewiczAndRaybourn2018}.  To improve this situation, we need to first
fully understand the current state of the practice for research software
development.

@@

This ``state of the practice for domain X'' exercise builds off of prior work on
measuring/assessing the state of software development practice in several
research domains.  We have updated the work that was done previously for domains
such as Geographic Information Systems \cite{SmithEtAl2018_arXivGIS}, Mesh
Generators \cite{SmithEtAl2016}, Seismology software \cite{SmithEtAl2018}, and
Statistical software for psychology \cite{SmithEtAl2018_StatSoft}.  Initial
tests of the new methodology have been done for medical image analysis software
\cite{Dong2021} and for Lattice Boltzmann Method (LBM) software
\cite{Michalski2021}.

@@

In the proposed methodology, the collected data is combined to rank the software
within each domain using the Analytic Hierarchy Process (AHP) \cite{Saaty1980}.
As in the previous measurement exercise, we use AHP to develop a list of
software ranked by quality.  However, in the new process we do not stop with
this list.  The Domain Expert is consulted to verify the ordering, and to
discuss the decisions that led to the ranking.  The AHP process is used to
facilitate a conversation with the Domain Expert as a means to deepen our
understanding of the software in the domain, and the needs of typical
developers.

@@

The scope of this methodology includes observations on product, artifact
(documentation, test scripts, etc.) and process quality for research software.
We leave the assessment of the performance of research software, for instance
using benchmarks, to other projects, such as the work of
\cite{kaagstrom1998gemm}.  Currently we are also leaving experiments to measure
usability and modifiability as future work, as discussed in
Section~\ref{SecFutureWork}.

@@

A domain of research software must be identified to begin the assessment.
Research software is defined in this exercise as ``software that is used to
generate, process or analyze results that [are intended] to appear in a
publication'' \cite{hettrick2014uk} in a scientific or engineering context.
Research software is a more general term for what is often called scientific
computing.  To be applicable for the methodology described in this document, the
chosen domain must have the following properties:

@@

Some examples of domains that fit these criteria are finite element analysis
\cite{szabo1996finite}, quantum chemistry \cite{veryazov20042molcas},
seismology \cite{SmithEtAl2018}, as well as mesh generators
\cite{smith2016state}.

@@

Since each domain we will study will have a reasonably small scope, we will be
able to view the software as constituting a program family.  The concept of a
program family is defined by \cite{parnas1976design} as ``a set of programs
whose common properties are so extensive that it is advantageous to study the
common properties of the programs before analyzing individual members''.
Studying the common properties within a family of related programs is termed a
domain analysis.

@@

\begin{figure}[!ht]
	\begin{center}
	  %\includegraphics[width=1.0\textwidth]{measurement_template}
	  \caption{Excerpt of the Top Section of the Measurement Template (Summary Information)}
	  \label{measurement_template_image}
	\end{center}
  \end{figure}

The full template consists of 108 questions categorized under 9 qualities.  The
questions were designed to be unambiguous, quantifiable and measurable with
limited time and domain knowledge. The measures are grouped under headings for
each quality, and one for summary information. The summary information (shown in
Figure~\ref{measurement_template_image}) is the first section of the template.
This section summarizes general information, such as the software name, number
of developers, etc.  We follow the definitions given by
\cite{GewaltigAndCannon2012} for the software categories.  Public means software
intended for public use.  Private means software aimed only at a specific
group, while the concept category is used for software written simply to demonstrate
algorithms or concepts. The three categories of development models are: open
source, where source code is freely available under an open source license;
free-ware, where a binary or executable is provided for free; and, commercial,
where the user must pay for the software product.

@@

As in \cite{SmithEtAl2016}, Virtual machines (VMs) are used to provide an
optimal testing environments for each package. VMs were used because it is
easier to start with a fresh environment without having to worry about existing
libraries and conflicts. Moreover, when the tests are complete the VM can be
deleted, without any impact on the host operating system. The most significant
advantage of using VMs is to level the playing field. Every software install
starts from a clean slate, which removes ``works-on-my-computer'' errors. When
filling in the measurement template spreadsheet, the the details for each VM
should be noted, including hypervisor and operating system version.

@@

In the previous state of the practice measurement process
\cite{SmithEtAl2018_arXivGIS, SmithEtAl2016, SmithEtAl2018}, we only based our
assessment on information available in on-line software repos.  However, this
approach meant we weren't able to learn about the development process, the
attitudes of the developers, the pain points for developers and how the
developers handle modifiability, reproducibility and usability.  Therefore, in
the reboot of the measurement process, we have explicitly added a stage for
interviewing research software developers.

@@

Table~\ref{TabPersonHours} estimates the time required (in person hours) to
complete a state of the practice assessment for domain X.  The table assumes
that the domain has already been decided and the Domain Expert has been
recruited.  The time spent by the Domain Expert is not included in the numbers
shown in the table, since the amount of time that the domain expert will work
independently of the rest of the assessment team will be small.  Moreover, this
amount of time will vary greatly depending on the preferred work habits of the
Domain Expert.  The table follows the steps outlined in Section~\ref{StepsAQDS}.
Time is not included for reviewing the methodology. Moreover, it is assumed that
the template spreadsheets linked in this document, and the developed AHP tool,
will be employed, rather than developing new tools.  The person hours given are
a rough estimate, based on our experience completing assessments for medical
image analysis software \cite{Dong2021} and for Lattice Boltzmann Method (LBM)
software \cite{Michalski2021}.  These two domains were assessed at the same
time as designing the methodology presented in this document.  We did our best
to estimate the time spent on measurement and separate it from the time spend on
design and development.  The estimate assumes 30 software packages will be
measured; the numbers will need to be adjusted if the total packages changes.

@@

As explained in the introduction (Section~\ref{SecIntroduction}), our eventual
goal is to improve the state of practice of software development, which requires
a baseline means for measuring the current state of the practice.  Now that we
have a methodology for assessing the state of practice for a given domain, the
next task is to complete the measurements for multiple domains.  Using the
previous and updated methodologies, we have measured the state of
practice for the following domains: Geographic Information Systems
\cite{SmithEtAl2018_arXivGIS}, Mesh Generators \cite{SmithEtAl2016},
Seismology software \cite{SmithEtAl2018}, Statistical software for psychology
\cite{SmithEtAl2018_StatSoft}, medical image analysis software \cite{Dong2021}
and for Lattice Boltzmann Method software \cite{Michalski2021}.  Future domains
for measurement could include finite element software, computational medicine,
machine learning, ordinary differential equation solvers, computer graphics,
stoichiometry, etc.

@@

If we can add measures for modifiability and usability, we can start to measure
the quality impact of software development processes, tools and techniques.  For
instance, we have a project (called
\href{https://github.com/JacquesCarette/Drasil} {Drasil}
\cite{SzymczakEtAl2016}), which facilitates a development process that focuses
on knowledge capture, followed by generation of code and other software
artifacts from the captured knowledge.  To understand the advantages and
disadvantages of Drasil, we could measure its quality.  In particular, we would
like to see the impact of a generative process on the quality of usability and
modifiability.

@@

The next experiment is designed to gather qualitative data regarding the
modifiability of each software package. This proposed experiment also involves
experimental subjects/participants, who in this case are asked to modify a set
of software packages.  The specific modifications requested will again depend on
the software domain.  In advance of the experiment the Domain Expert will be
asked the likely changes for software in this domain.  We emphasize likely
changes, instead of any changes because software cannot be designed so that
everything is equally easy to change \cite{parnas1986rational}.  The procedure
could be along the following lines:

@@

Combining the quantitative data from the measurement template with the interview
results, along with the domain experts knowledge, we can determine the current
state of the practice for domain X.  Using our methodology, spreadsheet
templates and AHP tool, we estimate (based on our experience with using the
process) the time to complete an assessment for a given domain at 173 person
hours.

\subsubsection{Acknowledgements} Please place your acknowledgments at
the end of the paper, preceded by an unnumbered run-in heading (i.e.
3rd-level heading).

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}
